---
title: "Experiment 1 figures"
author: "English"
date: "7/6/2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction

Plastic agricultural mulches provide weed control, moisture retention, temperature optimization, and decreased erosion. Because conventional mulches are composed of polyethylene, a poorly biodegradable plastic, they often end up in landfills after harvest. A more sustainable alternative is the use of biodegradable plastic mulches (BDMs) which, after being tilled into the soil subsequent to service life, degrade into carbon dioxide, water, and soil organic matter.  


However, the role of temperature in the biodegradation of these materials is not well known. We conducted a manipulative incubation experiment to look at in situ biodegradability of a mulch films: a starch/polyester blend (BioAgri). They were kept at different temperatures (10, 20, and 30°C) and the evolution of CO2 was measured biweekly over 16 weeks to monitor the time course of biodegradation. The influence of temperature on the rate of degradation for biodegradable mulches could help determine the impact of climate and environment on microbial assimilation.

# Methods 

CO2 evolution was measured every 2 weeks by sampling the headspace and measuring CO2 concentration on an infrared gas analyzer. 
At t=0 and t=16 weeks total organic carbon (TOC) and bulk soil carbon isotopic signature were measured using Cavity Ring-Down Spectroscopy and microbial biomass and EOC were measured using the chloroform extraction method.


# Results 

<br>
  <br>

### DOC 

```{r echo= FALSE, message= FALSE, warning= FALSE}

library(tidyverse)
library(readr)
library (ggplot2)
library (knitr)

EXP1DOC<-read_csv("../data/exp1docmbc.csv")

#DOC table
EXP1DOC_2<- EXP1DOC %>%
  group_by (Mulch,Temp)  %>%
  summarise (avg = mean (DOC), n = length(DOC), sd = sd(DOC), 
             se = sd/sqrt(n), ymax = avg +se, ymin = avg -se)

kable(EXP1DOC_2)



#4 plots of DOC (each mulch with each column a different temperature)
EXP1DOC_plot<- EXP1DOC_2 %>%
  ggplot(aes(x= Temp, y= avg)) + 
  geom_bar(stat = "identity", fill="dodgerblue3") + 
  geom_errorbar((aes(ymax = ymax, ymin= ymin)), width=0.2) +
  ylab("DOC ug-C/ g-Dry soil\n")+ xlab("\nTemperature") +
  theme (
    axis.text = element_text(size = 8)) + 
  facet_wrap(~Mulch)
plot(EXP1DOC_plot)

```
<br>
  <br>

### MBC

```{r echo= FALSE, message= FALSE, warning= FALSE}

#MBC table
EXP1MBC_2<- EXP1DOC %>%
  group_by (Mulch,Temp)  %>%
  summarise (avg = mean (MBC), n = length(MBC), sd = sd(MBC), 
             se = sd/sqrt(n), ymax = avg +se, ymin = avg -se)
kable(EXP1MBC_2)

#MBC plot
EXP1MBC_plot<- EXP1MBC_2 %>%
  ggplot(aes(x= Temp, y= avg)) + 
  geom_bar(stat = "identity", fill="dodgerblue") + 
  geom_errorbar((aes(ymax = ymax, ymin= ymin)), width=0.2) +
  ylab("MBC ug-C/ g-Dry soil\n")+ xlab("\nTemperature") +
  theme (
    axis.text = element_text(size = 8)) + 
  facet_wrap(~Mulch)
EXP1MBC_plot


```
<br>
  <br>

### TOC

```{r echo= FALSE, message= FALSE, warning= FALSE}
EXP1TOC<-read_csv("../data/exp1tocd13c.csv")

EXP1TOC$Temp <- as.character(EXP1TOC$Temp)

#TOC table
EXP1TOC_2<- EXP1TOC %>%
  group_by (Mulch,Temp)  %>%
  summarise (avg = mean (TOC), n = length(TOC), sd = sd(TOC), 
             se = sd/sqrt(n), ymax = avg +se, ymin = avg -se)
kable(EXP1TOC_2) 

#1 plots of TOC 
EXP1TOC_plot<- EXP1TOC_2 %>%
  ggplot(aes(x= Temp, y= avg)) + 
  geom_bar(stat = "identity", fill="coral1") + 
  geom_errorbar((aes(ymax = ymax, ymin= ymin)), width=0.2) +
  ylab("TOC mg-C/ g-Dry soil\n")+ xlab("\nTemperature") +
  theme (
    axis.text = element_text(size = 8))+
  facet_wrap(~Mulch)
EXP1TOC_plot



```
<br>
  <br>
  
### d13C 

```{r echo= FALSE, message= FALSE, warning= FALSE}

EXP1TOC<-read_csv("../data/exp1tocd13c.csv")

EXP1TOC$Temp <- as.character(EXP1TOC$Temp)

#d13C table
EXP1d13C_2<- EXP1TOC %>%
  group_by (Mulch,Temp)  %>%
  summarise (avg = mean (d13C), n = length(d13C), sd = sd(d13C), 
             se = sd/sqrt(n), ymax = avg +se, ymin = avg -se)
kable(EXP1d13C_2)

EXP1d13C_plot<- EXP1d13C_2 %>%
  ggplot(aes(x= Mulch, y= avg, colour = Temp)) + 
  geom_point () +
  #geom_bar(stat = "identity", fill="coral3") + 
  geom_errorbar((aes(ymax = ymax, ymin= ymin)), width=0.2) +
  ylab("d13C\n")+ xlab("\nTreatment") +
  scale_fill_discrete() +
  theme (
    axis.text = element_text(size = 8)) 
EXP1d13C_plot



```

<br>
  <br>
  
### CO2 respiration 

```{r echo= FALSE, message= FALSE, warning= FALSE}
co2evol<-read_csv("../data/redoco2evolved4.csv", col_types = cols(
  meday = col_date("%m/%d/%y")))

#30C 
co2evol_30<- co2evol %>%
  group_by (memulch, metemp, meday)  %>%
  filter (metemp == "30") %>%
  summarise (avg = mean (meco2), n = length(meco2), sd = sd(meco2), 
             se = sd/sqrt(n), ymax = avg +se, ymin = avg -se) %>%
  ggplot(aes(meday,avg, colour = memulch)) + 
  geom_point() + geom_line() +
  geom_errorbar((aes(ymax = ymax, ymin= ymin)), width=0.2) +
  ylim (0, 1250) +
  ylab(expression("Cumulative CO"[2]*~mu*"g-C/g-Dry soil"))+ xlab("\nDate") +
  scale_color_discrete(name = "Treatment", labels = c("BioAgri","Cellulose", "No mulch", "PLA/PHA")) + 
  theme (
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size= 12),
    axis.title.x = element_text(size= 14),  
    axis.title.y = element_text(size= 14), 
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 12))
co2evol_30


#20C 
co2evol_20<- co2evol %>%
  group_by (memulch, metemp, meday)  %>%
  filter (metemp == "20") %>%
  summarise (avg = mean (meco2), n = length(meco2), sd = sd(meco2), 
             se = sd/sqrt(n), ymax = avg +se, ymin = avg -se) %>%
  ggplot(aes(meday,avg, colour = memulch)) + 
  geom_point() + geom_line() +
  geom_errorbar((aes(ymax = ymax, ymin= ymin)), width=0.2) +
  ylim (0, 1250) +
  ylab(expression("Cumulative CO"[2]*~mu*"g-C/g-Dry soil"))+ xlab("\nDate") +
  scale_color_discrete(name = "Treatment", labels = c("BioAgri","Cellulose", "No mulch", "PLA/PHA")) +
  theme (
      axis.text.x = element_text(size = 12),
      axis.text.y = element_text(size= 12),
      axis.title.x = element_text(size= 14),  
      axis.title.y = element_text(size= 14), 
      legend.title = element_text(size = 12),
      legend.text = element_text(size = 12))
co2evol_20

co2evol_10<- co2evol %>%
  group_by (memulch, metemp, meday)  %>%
  filter (metemp == "10") %>%
  summarise (avg = mean (meco2), n = length(meco2), sd = sd(meco2), 
             se = sd/sqrt(n), ymax = avg +se, ymin = avg -se) %>%
  ggplot(aes(meday,avg, colour = memulch)) + 
  geom_point() + geom_line() +
  geom_errorbar((aes(ymax = ymax, ymin= ymin)), width=0.2) +
  ylab(expression("Cumulative CO"[2]*~mu*"g-C/g-Dry soil"))+ xlab("\nDate") +
  scale_color_discrete(name = "Treatment", labels = c("BioAgri","Cellulose", "No mulch", "PLA/PHA")) +
    theme (
      axis.text.x = element_text(size = 12),
      axis.text.y = element_text(size= 12),
      axis.title.x = element_text(size= 14),  
      axis.title.y = element_text(size= 14), 
      legend.title = element_text(size = 12),
      legend.text = element_text(size = 12))
co2evol_10


```
<br>
  <br>
  
### Enzymes

```{r echo= FALSE, message= FALSE, warning= FALSE}
EXP1ENZ<-read_csv("../data/exp1enz.csv")

EXP1ENZ$Temp <- as.character(EXP1ENZ$Temp)

#Enzyme table
EXP1ENZ_2<- EXP1ENZ %>%
  group_by (Mulch,Temp, Enzyme)  %>%
  summarise (avg = mean (Activity), n = length(Activity), sd = sd(Activity), 
             se = sd/sqrt(n), ymax = avg +se, ymin = avg -se)

#1 plots of ENZYMES 
EXP1ENZ_plot<- EXP1ENZ_2 %>%
  ggplot(aes(x= Mulch, y= avg, fill = Temp)) + 
  geom_bar(stat = "identity", position=position_dodge()) + 
  geom_errorbar((aes(ymax = ymax, ymin= ymin)), position=position_dodge(0.9), width=0.2) +
  ylab("Activity\n")+ xlab("\nMulch") +
  theme (
    axis.text = element_text(size = 8))+
  facet_wrap(~Enzyme, scales = "free_y")
EXP1ENZ_plot

```

